### 2014-08-11 Docs/flowchart/flowchart.tex
more equationing

The equation of state is cast into an elliptic constraint on the
velocity field by differentiating $p_0(\rho, s, X_k)$ along particle
paths, giving:
\begin{equation}
\nabla \cdot (\beta_0 \Ub) = 
   \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
\label{eq:U divergence}
\end{equation}
where $\beta_0$ is a density-like variable that carries background
 stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
and 
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext
\end{equation}
where $p_{X_k} \equiv \left. \partial p / \partial X_k
\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
$c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.

@@

In practice, we calculate $w_0$ by integrating
the one-dimensional divergence constraint:
\begin{equation}
\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
\end{equation}
Then we define
\begin{equation}
- \frac{1}{\rho_0} \frac{\partial \pizero}{\partial r} = \frac{\partial w_0}{\partial t} + 
   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}
once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
Then we can include this for completeness in the update for $\ut.$

@@

We decompose the full velocity field into a base state velocity,
$w_0$, that governs the base state dynamics, and a local velocity,
$\Ubt$, that governs the local dynamics, i.e., 
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$.
The velocity evolution equations are then \MarginPar{we need the $\beta_0$ here}
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0
  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
  - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = 
    \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}
           \frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting
(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

@@

>>>>

The equation of state is cast into an elliptic constraint on the
velocity field by differentiating $p_0(\rho, s, X_k)$ along particle
paths, giving:
\begin{equation}
\nabla \cdot (\beta_0 \Ub) = 
   \beta_0 \left ( S - \frac{1}{\gammabar p_0} \frac{\partial p_0}{\partial t} \right )
\label{eq:U divergence}
\end{equation}
where $\beta_0$ is a density-like variable that carries background
 stratification, defined as
\begin{equation}
\beta_0(r,t) = \rho_0(0,t)\exp\left(\int_0^r\frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial r'}dr'\right),
\end{equation}
and 
\begin{equation}
S = -\sigma\sum_k\xi_k\omegadot_k + \frac{1}{\rho p_\rho}\sum_k p_{X_k}\omegadot_k + \sigma\Hnuc + \sigma\Hext + \frac{\sigma}{\rho} \nabla \cdot \kth \nabla T
\end{equation}
where $p_{X_k} \equiv \left. \partial p / \partial X_k
\right|_{\rho,T,X_{j,j\ne k}}$, $\xi_k \equiv \left. \partial h / 
\partial X_k \right |_{p,T,X_{j,j\ne k}}, 
p_\rho \equiv \left. \partial p/\partial \rho \right |_{T, X_k}$, and
$\sigma \equiv p_T/(\rho c_p p_\rho)$, with
$p_T \equiv \left. \partial p / \partial T \right|_{\rho, X_k}$ and
$c_p \equiv \left.  \partial h / \partial T
\right|_{p,X_k}$ is the specific heat at constant pressure.   The last
term is only present if we are using thermal diffusion (\runparam{use\_thermal\_diffusion} {\tt = T}).  In this term, $\kth$ is the thermal conductivity.

@@

@@

We decompose the full velocity field into a base state velocity,
$w_0$, that governs the base state dynamics, and a local velocity,
$\Ubt$, that governs the local dynamics, i.e., 
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$---the motivation for this splitting was given in paper~II.
The velocity evolution equations are then 
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r},\label{eq:w0
  evolution}\\ 
%
\frac{\partial\Ubt}{\partial t} &=& -(\Ubt + w_0\er)\cdot\nabla\Ubt
  - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{\beta_0}{\rho_0}\frac{\partial(\pi_0/\beta_0)}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = 
    \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}
           \frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting
(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

\subsection{Base State Expansion}

In practice, we calculate $w_0$ by integrating
the one-dimensional divergence constraint:
\begin{equation}
\frac{\partial w_0}{\partial r} = \Sbar - \frac{1}{\gammabar p_0} \etarho g \enskip 
\end{equation}
Then we define \MarginPar{this term needs the $\beta_0$'s as shown.  Also, is this really
the form, or should it be $\overline{(\beta_0/\rho_0) \partial (\pi_0/\beta_0)/\partial r}$?}
\begin{equation}
- \frac{\beta_0}{\rho_0} \frac{\partial (\pizero/\beta_0)}{\partial r} = \frac{\partial w_0}{\partial t} + 
   w_0 \frac{\partial w_0}{\partial r} \enskip, \label{eq:pizero}
\end{equation}
once $w_0$ at the old and new times is known, and the advective term is computed explicitly.
Then we can include this for completeness in the update for $\ut.$

@@

In paper~III, we showed that for a plane-parallel atmosphere with constant gravity,
$\psi = \etarho g$
